---
title: "Bar Graphs"
author: "Isaac Neuhaus"
date: "06/04/2024"
output:
  html_document:
    toc: yes
    theme: cerulean
    smart: no
    self_contained: yes
editor_options: 
  chunk_output_type: console
---

# Load data and Libraries
### https://r4ds.had.co.nz/data-visualisation.html
```{r setup}
if (!("ggplot2" %in% installed.packages())) {
    install.packages('ggplot2')
}
library('ggplot2')
if (!("devtools" %in% installed.packages())) {
    install.packages('devtools')
}
if (!("canvasXpress" %in% installed.packages())) {
    devtools::install_github('neuhausi/canvasXpress')
}
#devtools::install_local("~/git/canvas/R/canvasXpress.tar.gz", build_manual = TRUE, upgrade = "always")
library('canvasXpress')
```


## geom_bar is designed to make it easy to create bar charts that show counts (or sums of weights)
```{r}
g <- ggplot(mpg, aes(class)) + geom_bar()
g
canvasXpress(g)
```

## Total engine displacement of each class
```{r}
g <- ggplot(mpg, aes(class)) + geom_bar(aes(weight = displ))
g
canvasXpress(g)
```

## Map class to y instead to flip the orientation
```{r}
g <- ggplot(mpg) + geom_bar(aes(y = class))
g
canvasXpress(g)
```

## Bar charts are automatically stacked when multiple bars are placed at the same location. The order of the fill is designed to match the legend
```{r}
g <- ggplot(mpg, aes(class)) + geom_bar(aes(fill = drv))
g
canvasXpress(g)
```

## If you need to flip the order (because you've flipped the orientation) call position_stack() explicitly
```{r}
g <- ggplot(mpg, aes(y = class)) + geom_bar(aes(fill = drv), position = position_stack(reverse = TRUE)) + theme(legend.position = "top")
g
canvasXpress(g)
```

## Statistical transformations
```{r}
sample_data <- diamonds[sample(nrow(diamonds), 100), ]
g = ggplot(data = sample_data) + geom_bar(mapping = aes(x = cut))
g
canvasXpress(g)
```

## Position adjustments - No Position
```{r}
sample_data <- diamonds[sample(nrow(diamonds), 100), ]
g = ggplot(data = sample_data) + geom_bar(mapping = aes(x = cut, fill = cut))
g
canvasXpress(g)
```

## Position adjustments - No Position 2
```{r}
sample_data <- diamonds[sample(nrow(diamonds), 100), ]
g = ggplot(data = sample_data) + geom_bar(mapping = aes(x = cut, fill = clarity))
g
canvasXpress(g)
```

## Position adjustments - Fill
```{r}
sample_data <- diamonds[sample(nrow(diamonds), 100), ]
g = ggplot(data = sample_data) + geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")
g
canvasXpress(g)
```

## Position adjustments - Dodge
```{r}
sample_data <- diamonds[sample(nrow(diamonds), 100), ]
g = ggplot(data = sample_data) + geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
g
canvasXpress(g)
```

## Special Fill
```{r}
sample_data <- diamonds[sample(nrow(diamonds), 5), ]
g = ggplot(data = sample_data, aes(x = 1, y = price, fill = cut)) + geom_bar(position="fill", stat="identity")
g
canvasXpress(g)
```

## Special Fill 2
```{r}
sample_data <- diamonds[sample(nrow(diamonds), 50), ]
g = ggplot(data = sample_data, aes(x = 1, y = price, fill = cut)) + geom_bar(position="fill", stat="identity")
g
canvasXpress(g)
```
