---
title: "CanvasXpress GGPlot Demo"
author: "Isaac Neuhaus"
date: "02/04/2025"
output:
  html_document:
    toc: yes
    theme: cerulean
    smart: no
    self_contained: yes
---

### Install and load the necessary packages

```{r}
if (!("ggplot2" %in% installed.packages())) {
    install.packages('ggplot2')
}
library('ggplot2')
if (!("canvasXpress" %in% installed.packages())) {
    devtools::install_github('neuhausi/canvasXpress')
}
library('canvasXpress')
```

This demo shows the ability to easily integrate CanvasXpress with ggplot in R.

### Data: mpg dataset

This dataset contains a subset of the fuel economy data that the EPA makes available on <https://fueleconomy.gov/>. It contains only models which had a new release every year between 1999 and 2008 - this was used as a proxy for the popularity of the car.

```{r}
head(mpg)
```

### Plot with ggplot

Compare displacement (displ) vs highway efficiency (hwy) and color the data by vehicle class (class). Store the ggplot object in a variable.

```{r}
s = ggplot(mpg, aes(displ, hwy)) + geom_point(aes(color = class))
s
```

### Plot with CanvasXpress

Use the ggplot object stored in the variable 's' and plot with CanvasXpress

```{r}
canvasXpress(s)
```

### Use the native CanvasXpress API to plot cty vs hwy, color by class

```{r}
canvasXpress(mpg, type = "scatter2d", xAxis = "cty", yAxis = "hwy", colorBy = "class", theme = "ggplot")
```

### Create a visualization with plain English

```{r}
canvasXpress(mpg, 
  prompt = "Create a Scatter 2d plot of cty vs hwy, 
            color by class. Use the ggplot style."
)
```
