---
title: "CanvasXpress for Python - Rmarkdown"
author: "Todd Brett"
date: "06/27/2023"
output:
  html_document:
    toc: yes
    theme: cerulean
    smart: no
    self_contained: yes
---

There are multiple options for activating Python kernals in RMD documents, but the most convenient method is to
activate reticulate.  

```{r}
library('reticulate')
```

The CanvasXpress for Python (CX-P) package needs to be installed on the host system.  Python packages are installed
into the active Python environment, so a virtual environment (a copy of a Python environment for isolated use) is
recommended to keep the global installation clean.

The following bash command, for Linux/UNIX/MACOS systems, checks for a local virtual environment; if no such environment
is found then a new one is created using the system Python3 installation.

```{bash, engine.opts='-l'}
if ! [[ -d ./venv ]] ; then
    python3 -m venv ./venv
fi
```

reticulate is then instructed to use the new virtual environment.  CX-P, along with its required dependencies, 
will now available for general use.

```{r}
use_virtualenv('./venv', required=TRUE)
py_install('canvasxpress[all]')
```

Data and configurations are defined as normal, such as via a `dict`.

```{python}
demo_data = {
  "y": {
    "vars": ["V1"],
    "smps": ["S1","S2","S3"],
    "data": [
      [33,44,55]
    ]
  }
}

demo_config = {
  "graphOrientation": "vertical",
  "plotBox": True,
  "showLegend": False,
  "smpLabelRotate": 90,
  "smpTitle": "Samples",
  "theme": "CanvasXpress",
  "title": "Bar Graph Title",
  "xAxis": ["V1"],
  "xAxisTitle": "Value",
  "graphType": "Bar"
}
```

Charts can reuse pre-declared data and configurations, such as those above, or they can use new data and configurations 
declared at the time the object is constructed.

```{python}
from canvasxpress.canvas import CanvasXpress

example1_chart = CanvasXpress(
  render_to = "Example1",
  data = demo_data,
  config = demo_config,
)

```

Chart objects are converted into HTML (etc.) via a render object.  Render objects are initialized with a set of charts,
and at the time the HTML is to be generated the `render` method is called.  The render object to be used depends on the
environment in which the chart will be created.

```{python}
from canvasxpress.render.rstudio import CXHtml

renderer = CXHtml(example1_chart)
renderer.render()
```

CX-P provides full and interactive integration with contexts such as Flask, Dash, and similar; however, for RMD at this
time only HTML is generated.  A Shiny for Python solution is being developed, but until then an IPython widget is provided
to the RMD for use with Pandoc.  If only the chunk is executed then an IPython object is noted, but no visualization will
be provided as RMD has yet to fully integrated; however, Pandoc does support IPython widgets and upon being knitted the
RMD will include a live CanvasXpress chart from the Python chunk.

Shiny for R does support the Python-generated HTML, and as such if the HTML data is obtained it can be displayed in RStudio
once the combined Python and R chunks are executed.  Shiny will create an RMD-compliant HTML container widget and load the 
provided HTML for immediate display.

```{python}
example2_chart = CanvasXpress(
  render_to = "Example2",
  data = demo_data,
  config = demo_config,
)

renderer = CXHtml(example2_chart)
cx2_html = renderer.render().data
```

```{r}
shiny::HTML(py$cx2_html)
```

Multiple charts can be rendered via a code chunk, and in this case charts can be distributed in rows across a series
of columns.  By default a single column is used and multiple charts will stack.

Note that each chart `render_to` target must be unique on the page.  Anonymous charts can be created by setting the 
render_to property to an empty string, at which point the renderer will automatically assign GUIDs at the time of
render.

```{python}
renderer = CXHtml(
  [
    CanvasXpress(
      render_to = "",
      data = demo_data,
      config = demo_config,
    ),
    CanvasXpress(
      render_to = "",
      data = demo_data,
      config = demo_config,
    ),
    CanvasXpress(
      render_to = "",
      data = demo_data,
      config = demo_config,
    )
  ]
)
multi_chart_html = renderer.render(columns=2).data
```

```{r}
shiny::HTML(py$multi_chart_html)
```
