---
title: "CanvasXpress Demo Notebook"
author: "Data Table"
date: "07/03/2024"
output:
  html_document:
    toc: yes
    theme: cerulean
    smart: no
    self_contained: yes
---


```{r include=FALSE}
if (!("canvasXpress" %in% installed.packages())) {
    devtools::install_github('neuhausi/canvasXpress')
}
#devtools::install_local("~/git/canvas/R/canvasXpress.tar.gz", build_manual = TRUE, upgrade = "always")
library('canvasXpress')
```

```{r include=FALSE}
data <- read.table("https://www.canvasxpress.org/assets/documents/examples/Data/restaurant_customer_satisfaction.csv", header=TRUE, sep=",", quote="\"")
```

# Dataset: Restaurant Customer Satisfaction
#### (https://www.kaggle.com/datasets/rabieelkharoua/predict-restaurant-customer-satisfaction-dataset)

```{r echo=FALSE}
canvasXpress(data,
             view = 'table',
             dataTableTransposed = FALSE,
             colWidth = 150,
             width = 910, 
             height = 600)
```

# Componnets of the Data Table

* Scrollbar - Vertical and horizontal
* Search widget - Close icon
  + Show example
* Column header - Number of rows and columns
* Toolbar with icons
* Header row - Column names with dropdown button for sorting and selecting columns
  + Hide/ show columns
* Data rows - Data values

# Icons in the Toolbar

* Save button - Save the table as a csv file
* Explore metadata - View the metadata of the table
  + Click on some graphs to filter data
  + Press ESC to reset the filters
  + Remove "Explore metadata" and Show Data Table
* Filter data - Filter the data based on the column values (ala Spotfire)
  + Click or adjust filters to filter data
  + Click on reset filter icon to reset the filters
  + CLose the filter panel to hide the filters
* Maximize - Maximize the data table
  + Minimize the data table with the same icon
  + Don't close the data table - It will show the plot
    + Maximize and then minimize the plot and then click on the data table icon
* Color Data - Color the data based on the column values
  + Toggle the color data on and off
* Transpose - Transpose the data table
  + Toggle the transposed data table on and off
* Clear filters - Clear the filters applied to the data table

