---
title: "CanvasXpress Demo"
author: "Isaac Neuhaus"
date: "06/04/2024"
output:
  html_document:
    toc: yes
    theme: cerulean
    smart: no
    self_contained: yes
---

```{r setup, include=FALSE}
if (!("ggplot2" %in% installed.packages())) {
    install.packages('ggplot2')
}
library('ggplot2')
if (!("canvasXpress" %in% installed.packages())) {
    devtools::install_github('neuhausi/canvasXpress')
}
#devtools::install_local("~/git/canvas/R/canvasXpress.tar.gz", build_manual = TRUE, upgrade = "always")
library('canvasXpress')
```

This demo shows the ability to easily integrate CanvasXpress with ggplot in R. There are similar demos for python.

## Scatter2d Plot
### Data: mpg dataset

This dataset contains a subset of the fuel economy data that the EPA makes available on https://fueleconomy.gov/. It contains only models which had a new release every year between 1999 and 2008 - this was used as a proxy for the popularity of the car.

```{r}
head(mpg)
```

### Plot with ggplot

Compare displacement (displ) vs highway efficiency (hwy) and color the data by vehicle class (class). Store the ggplot object in a variable.

```{r}
s = ggplot(mpg, aes(displ, hwy)) + geom_point(aes(color = class))
```


### Plot with CanvasXpress

Use the ggplot object stored in the variable 's' and plot with CanvasXpress

```{r}
canvasXpress(s, width=800, height=600, shapeBy="drv", selectedDataPoints=list("1", "100", "101", "102", "103", "105", "194", "19"))
```
