---
title: "ggblanket"
author: "Isaac Neuhaus"
date: "06/04/2024"
output:
  html_document:
    toc: yes
    theme: cerulean
    smart: no
    self_contained: yes
---

### How it works

```{r}
if (!("canvasXpress" %in% installed.packages())) {
    devtools::install_github('neuhausi/canvasXpress')
}
#devtools::install_local("~/git/canvas/R/canvasXpress.tar.gz", build_manual = TRUE, upgrade = "always")
library(ggblanket)
library(ggplot2)
library(dplyr)
library(stringr)
library(tidyr)
library(palmerpenguins)
library(patchwork)
library(canvasXpress)

set_blanket()
```


### Almost forty gg_* wrapper functions

```{r}
p <- penguins |>
  gg_point(
    x = flipper_length_mm,
    y = body_mass_g,
  )
p
canvasXpress(p, width = 672, height = 500)
```

### A single col argument to colour and fill by a variable

```{r}
p <- penguins |>
  gg_point(
    x = flipper_length_mm,
    y = body_mass_g, 
    col = species,
  )
p
canvasXpress(p, width = 672, height = 500)
```


```{r}
p <- penguins |>
  drop_na(sex) |>
  mutate(across(sex, str_to_sentence)) |>
  gg_bar(
    position = "dodge",
    y = species, 
    col = sex,
    width = 0.75,
  )
p
canvasXpress(p, width = 672, height = 500)
```

### A facet argument to facet by a variable

```{r}
p <- penguins |>
  drop_na(sex) |> 
  mutate(across(sex, str_to_sentence)) |> 
  gg_histogram(
    x = flipper_length_mm,
    facet = species,
    binwidth = 2
  )
p
canvasXpress(p, width = 672, height = 500)
```

### A facet2 argument to facet by a 2nd variable

```{r}
p <- penguins |>
  mutate(across(sex, str_to_sentence)) |> 
  gg_histogram(
    x = flipper_length_mm,
    facet = species,
    facet2 = sex,
    binwidth = 2
  )
p
canvasXpress(p, width = 672, height = 500)
```

### Unspecified titles converted to snakecase::sentence_case

```{r}
p <- diamonds |>
  gg_hex(
    coord = coord_cartesian(clip = "on"), 
    x = carat,
    y = price,
    y_limits = c(0, 20000),
  )
p
canvasXpress(p, width = 672, height = 500)
```

### Prefixed arguments to customise x, y, col and facet properties

```{r}
p <- penguins |>
  drop_na(sex) |>  
  gg_jitter(
    x = species,
    y = body_mass_g,
    col = flipper_length_mm,
    facet = sex,
    x_labels = \(x) str_sub(x, 1, 1),
    y_breaks = scales::breaks_width(1000),
    y_expand_limits = 2000,
    y_labels = scales::label_number(big.mark = " "), 
    y_transform = "log10",
    y_title = "Body mass (g)",
    col_steps = TRUE,
    col_breaks = \(x) quantile(x, seq(0, 1, 0.25)),
    col_pal = viridisLite::rocket(n = 9, direction = -1),
    facet_labels = str_to_sentence,
  )
p
canvasXpress(p, width = 672, height = 500)
```

### Access to other geom_* arguments via ...

```{r}
p <- penguins |>
  mutate(across(sex, str_to_sentence)) |> 
  drop_na(sex) |> 
  gg_smooth(
    x = flipper_length_mm,
    y = body_mass_g,
    col = sex, 
    col_pal = c("#003f5c", "#ffa600"),
    colour = "#bc5090", 
    linewidth = 1, 
    linetype = "dashed",
    alpha = 1, 
    se = TRUE, 
    level = 0.999, 
  )
p
canvasXpress(p, width = 672, height = 500)
```

```{r}
p <- penguins |> 
  gg_boxplot(
    y = island, 
    x = flipper_length_mm, 
    col = species,
    fill = NA,
    position = position_dodge2(preserve = "single")
  ) 
p
canvasXpress(p, width = 672, height = 500)
```

### Access other aesthetics via mapping

```{r}
p <- penguins |> 
  gg_jitter(
    y = species, 
    x = flipper_length_mm, 
    col = species,
    mapping = aes(alpha = species, shape = species),
  ) +
  scale_alpha_manual(values = c(0.33, 1, 0.33)) +
  guides(alpha = guide_legend(reverse = TRUE))
p
canvasXpress(p, width = 672, height = 500)
```

### Families of *_mode_* themes that differ by legend placement

```{r}
p <- penguins |>
  gg_histogram(
    x = flipper_length_mm,
    col = species,
    title = "Penguin flipper length by species",
    subtitle = "Palmer Archipelago, Antarctica",
    caption = "Source: Gorman, 2020", 
    mode = light_mode_t(),
    binwidth = 2
  ) +
  labs(colour = NULL, fill = NULL)
p
canvasXpress(p, width = 672, height = 500)
```

```{r}
p <- penguins |>
  gg_histogram(
    x = flipper_length_mm,
    col = species,
    title = "Penguin flipper length by species",
    subtitle = "Palmer Archipelago, Antarctica",
    caption = "Source: Gorman, 2020", 
    mode = grey_mode_b(),
    binwidth = 2
  )
p
canvasXpress(p, width = 672, height = 500)
```

```{r}
p <- penguins |>
  gg_histogram(
    x = flipper_length_mm,
    col = species,
    title = "Penguin flipper length by species",
    subtitle = "Palmer Archipelago, Antarctica",
    caption = "Source: Gorman, 2020", 
    mode = dark_mode_r(),
    binwidth = 2
  )
p
canvasXpress(p, width = 672, height = 500)
```

### Set the mode and geom defaults with set_blanket

```{r}
set_blanket(grey_mode_r(), "#ffa600")
```

```{r}
p <- penguins |>
  gg_point(
    x = flipper_length_mm, 
    y = body_mass_g,
    x_breaks = scales::breaks_pretty(3),
  ) +
  geom_vline(xintercept = 200) +
  annotate("text", x = I(0.25), y = I(0.75), label = "Here")
p
canvasXpress(p, width = 672, height = 500)
```

```{r}
p <- penguins |> 
  gg_histogram(
    x = flipper_length_mm,
    x_breaks = scales::breaks_pretty(3),
    binwidth = 2
  ) +
  geom_vline(xintercept = 200) +
  annotate("text", x = I(0.75), y = I(0.75), label = "Here")
p
canvasXpress(p, width = 672, height = 500)
```

```{r}
set_blanket(dark_mode_r(), geom_colour = "#bc5090", annotate_colour = darkness[1])

p1 <- penguins |>
  gg_point(
    x = flipper_length_mm, 
    y = body_mass_g,
    x_breaks = scales::breaks_pretty(3),
  ) +
  geom_vline(xintercept = 200) +
  annotate("text", x = I(0.25), y = I(0.75), label = "Here")
p2 <- penguins |> 
  gg_histogram(
    x = flipper_length_mm,
    x_breaks = scales::breaks_pretty(3),
    binwidth = 2
  ) +
  geom_vline(xintercept = 200) +
  annotate("text", x = I(0.75), y = I(0.75), label = "Here")
p = p1 + p2
p
canvasXpress(p, width = 672, height = 500)
```

### No Facet
```{r}
set_blanket(colour = orange)

p1 <- penguins |>
  gg_point(
    x = flipper_length_mm,
    y = body_mass_g,
    col = island
  )
p
canvasXpress(p, width = 672, height = 500)


p <- penguins |>
  gg_bar(
    x = sex,
    col = island
  )
p
canvasXpress(p, width = 672, height = 500)

p <- penguins |>
  gg_histogram(
    x = flipper_length_mm,
    col = island,
    binwidth = 2
  )
p
canvasXpress(p, width = 672, height = 500)
```

### Facet
```{r}
set_blanket(colour = orange)

p <- penguins |>
  gg_point(
    x = flipper_length_mm,
    y = body_mass_g,
    col = island,
    facet = species
  )
p
canvasXpress(p, width = 672, height = 500)


p <- penguins |>
  gg_bar(
    x = sex,
    col = island,
    facet = species
  )
p
canvasXpress(p, width = 672, height = 500)
p <- penguins |>
  gg_histogram(
    x = flipper_length_mm,
    col = island,
    facet = species,
    binwidth = 2
  )
p
canvasXpress(p, width = 672, height = 500)
```


### Update Geom Defaults
```{r}
update_geom_defaults("boxplot",
                     aes(colour = "red", 
                         fill = "blue", 
                         alpha = 0.5, 
                         linewidth = 0.66)
                     )
p <- penguins |> 
  ggplot(aes(x = species, y = body_mass_g)) +
  geom_boxplot() 
p
canvasXpress::canvasXpress(p)
```


### Senssion Info
```{r}
sessionInfo()
```