---
title: "JavaScript - R-Studio"
author: "Isaac Neuhaus"
date: "06/01/2023"
output:
  html_document:
    toc: yes
    theme: cerulean
    smart: no
    self_contained: yes
---

<style>
  .noCanvas {
    clear: both;
    display: inline-block;
  }
  pre {
    clear: both;
  }
</style>

```{r setup}
if (!("ggplot2" %in% installed.packages())) {
    install.packages('ggplot2')
}
library('ggplot2')
if (!("jsonlite" %in% installed.packages())) {
    install.packages('jsonlite')
}
library('jsonlite')
if (!("canvasXpress" %in% installed.packages())) {
    install.packages('canvasXpress')
}
#devtools::install_local("~/git/canvas/R/canvasXpress.tar.gz")
#devtools::install_github('neuhausi/canvasXpress')
library('canvasXpress')
```

## CanvasXpress can be run in R-Studio in a very flexible manner

This document shows how to run CanvasXpress in many environments using different languages 

## canvasXpress library

Using the canvasXpress library developed for htmlwidgets

```{r}
library(canvasXpress)
vals = c(33,44,55)
vars = c("V1")
smps = c("S1","S2","S3")
data = as.data.frame(matrix(vals, nrow = 1, ncol = 3, byrow = TRUE, dimnames = list(vars, smps)))
canvasXpress(
  data = data,
  graphOrientation = "vertical",
  graphType = "Bar",
  showLegend = FALSE,
  smpLabelRotate = 90,
  smpTitle = "Samples",
  theme = "CanvasXpress",
  title = "Bar Graph Title",
  xAxisTitle = "Value"
)
```

## ggplot

CanvasXpress can be used with the ggplot object itself

```{r}
g <- ggplot(mpg, aes(class)) + geom_bar()
canvasXpress(g)
```

## JavaScript

Plain JavaScript: You need to include a canvas DOM element in the document before you call the {js} snipet bellow.
In this example the file contain the following HTML snippet.

```html
<div class="canvasXpress noCanvas">
  <canvas id="canvasPlainJS" width="613" height="613"></canvas>
</div>
```

<div class="canvasXpress noCanvas">
  <canvas id="canvasPlainJS" width="613" height="613"></canvas>
</div>

```{js}
var data = {
  "y" : {
    "data" : [
      [10,100,70,130,60]
    ],
    "smps" : ["S1","S2","S3","S4","S5"],
    "vars" : ["A"]
  }
}

var config = {
  "colorScheme":"CanvasXpress",
  "graphOrientation":"vertical",
  "graphType":"Area",
  "lineType":"rect",
  "objectBorderColor":"false",
  "showLegend":"false",
  "showSampleNames":"true",
  "showTransition":"false",
  "smpLabelRotate":"90",
  "smpTitle":"time"
}

var cX = new CanvasXpress("canvasPlainJS", data, config);
```

## JavaScript with R objects

Same as above but with an R object. In this example I convert a list into JSON and use it for the data.

```{r}
json = toJSON(list("y" = list("data" = list(c(55,44,33)),"smps"= c("S1","S2","S3"), "vars" = c("V1"))))
```

<script>var data = `r json`</script>

The following HTML snippet needs to be added in the document to make it available to the JavaScript code below.
Note that the quotes need to be backticks "`".

```html
<script>var data = 'r json'</script>
```

Similarily you need to add the following HTML snippet.

```html
<div class="canvasXpress noCanvas">
  <canvas id="canvasRJS" width="613" height="613"></canvas>
</div>
```

<div class="canvasXpress noCanvas">
  <canvas id="canvasRJS" width="613" height="613"></canvas>
</div>

```{js}

var config = {
  "graphOrientation":"vertical",
  "graphType":"Bar",
  "showLegend":"false",
  "smpLabelRotate":"90",
  "smpTitle":"Samples",
  "theme":"CanvasXpress",
  "title":"Bar Graph Title",
  "xAxisTitle":"Value"
}
  
var cX = new CanvasXpress("canvasRJS", data, config);
```

## cxplot

We convert the mpg data set into a two dimensional table with rownames and colname and subsequently to JSON

```{r}
d = mpg
nd = data.frame(lapply(d, as.character), stringsAsFactors = FALSE)
nd = tibble::add_column(nd, Id = row.names(d), .before = 1)
nd = tibble::add_row(nd, .before = 1)
nd[1,] = colnames(nd)
mpgJSON = toJSON(as.matrix(nd))
```

<script>var mpgJSON = `r mpgJSON`</script>

The following HTML snippet needs to be added in the document to make it available to the JavaScript code below.
Note that the quotes need to be backticks "`".

```html
<script>var mpgJSON = 'r mpgJSON'</script>
```

Similarily you need to add the following HTML snippet.

```html
<div class="canvasXpress noCanvas">
  <canvas id="canvasCXPlot" width="613" height="613"></canvas>
</div>
```

<div class="canvasXpress noCanvas">
  <canvas id="canvasCXPlot" width="613" height="613"></canvas>
</div>

```{js}
 var cxp = new cxplot("canvasCXPlot", mpgJSON, aes({"x":"displ","y":"hwy"}));
 cxp.geom_point({"color":"class"});
 cxp.gg_labels({"xAxisTitle":"displ","yAxisTitle":"hwy"});
```

## Python

Similarily you need to add the following HTML snippet.

```html
<div class="canvasXpress noCanvas">
  <canvas id="canvasPython" width="613" height="613"></canvas>
</div>
```

<div class="canvasXpress noCanvas">
  <canvas id="canvasPython" width="613" height="613"></canvas>
</div>

```{python}
from canvasxpress.canvas import CanvasXpress
from canvasxpress.render.jupyter import CXNoteBook

cx = CanvasXpress(
  render_to = "canvasPython",
  data = {
    "y": {
      "vars": ["V1"],
      "smps": ["S1","S2","S3"],
      "data": [
        [33,44,55]
      ]
    }
  },
  config = {
    "graphOrientation": "vertical",
    "plotBox": True,
    "showLegend": False,
    "smpLabelRotate": 90,
    "smpTitle": "Samples",
    "theme": "CanvasXpress",
    "title": "Bar Graph Title",
    "xAxis": ["V1"],
    "xAxisTitle": "Value",
    "graphType": "Bar"
  },
  width = 609,
  height = 609
)
 
display = CXNoteBook(cx)
display.render()
```
