---
title: "R for Data Science"
author: "Isaac Neuhaus"
date: "06/04/2024"
output:
  html_document:
    toc: yes
    theme: cerulean
    smart: no
    self_contained: yes
---

# Load data and Libraries
### https://r4ds.had.co.nz/data-visualisation.html
```{r setup}
if (!("ggplot2" %in% installed.packages())) {
    install.packages('ggplot2')
}
library('ggplot2')
if (!("tidyverse" %in% installed.packages())) {
    install.packages('tidyverse')
}
library('tidyverse')
if (!("dplyr" %in% installed.packages())) {
    install.packages('dplyr')
}
library('dplyr')
if (!("devtools" %in% installed.packages())) {
    install.packages('devtools')
}
if (!("canvasXpress" %in% installed.packages())) {
    devtools::install_github('neuhausi/canvasXpress')
}
#devtools::install_local("~/git/canvas/R/canvasXpress.tar.gz", build_manual = TRUE, upgrade = "always")
library('canvasXpress')
```

## Creating a ggplot
```{r}
g = ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy))
g
canvasXpress(g, width = 672, height = 500)
```

## Aesthetic mappings - Color
```{r}
g = ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = class))
g
canvasXpress(g, width = 672, height = 500)
```

## Size
```{r}
g = ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, size = class))
g
canvasXpress(g, width = 672, height = 500)
```

## Alpha
```{r}
g = ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
g
canvasXpress(g, width = 672, height = 500)
```

## Shape
```{r}
g = ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, shape = class))
g
canvasXpress(g, width = 672, height = 500)
```

## Facets - wrap
```{r}
g = ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_wrap(~ class, nrow = 2)
g
canvasXpress(g, width = 672, height = 500)
```

## Facets - grid
```{r}
g = ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_grid(drv ~ cyl)
g
canvasXpress(g, width = 672, height = 500)
```

## Geometric objects
```{r}
g = ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy))
g
canvasXpress(g, width = 672, height = 500)
```

## Smooth
```{r}
g = ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy))
g
canvasXpress(g, width = 672, height = 500)
```

## Smooth - Linetype
```{r}
g = ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
g
canvasXpress(g, width = 672, height = 500)
```

## Smooth - Color
```{r}
g = ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy, color = drv), show.legend = FALSE)
g
canvasXpress(g, width = 672, height = 500)
```

## Smoth  - Point Color
```{r}
g = ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + geom_smooth(mapping = aes(x = displ, y = hwy))
g
canvasXpress(g, width = 672, height = 500)
```

## Smooth - Color2
```{r}
g = ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + geom_point(mapping = aes(color = class)) + geom_smooth()
g
canvasXpress(g, width = 672, height = 500)
```

## Statistical transformations
```{r}
sample_data <- diamonds[sample(nrow(diamonds), 1000), ]
g = ggplot(data = sample_data) + geom_bar(mapping = aes(x = cut))
g
canvasXpress(g, width = 672, height = 500)
```

## Position adjustments - No Position
```{r}
sample_data <- diamonds[sample(nrow(diamonds), 1000), ]
g = ggplot(data = sample_data) + geom_bar(mapping = aes(x = cut, fill = cut))
g
canvasXpress(g, width = 672, height = 500)
```

## Position adjustments - No Position 2
```{r}
sample_data <- diamonds[sample(nrow(diamonds), 1000), ]
g = ggplot(data = sample_data) + geom_bar(mapping = aes(x = cut, fill = clarity))
g
canvasXpress(g, width = 672, height = 500)
```

## Position adjustments - Fill
```{r}
sample_data <- diamonds[sample(nrow(diamonds), 1000), ]
g = ggplot(data = sample_data) + geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")
g
canvasXpress(g, width = 672, height = 500)
```

## Position adjustments - Dodge
```{r}
sample_data <- diamonds[sample(nrow(diamonds), 1000), ]
g = ggplot(data = sample_data) + geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
g
canvasXpress(g, width = 672, height = 500)
```

## Coordinate systems
```{r}
g = ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + geom_boxplot()
g
canvasXpress(g, width = 672, height = 500)
```


## Vertical Error Bars
```{r}
df <- ToothGrowth
df$dose <- as.factor(df$dose)
df.summary <- df %>%
  group_by(dose) %>%
  summarise(
    sd = sd(len, na.rm = TRUE),
    len = mean(len)
  )
f <- ggplot(
  df.summary, 
  aes(x = dose, y = len, ymin = len-sd, ymax = len+sd)
  )
v <- f + geom_errorbar(width = 0.2) + geom_point(size = 1.5)
v
canvasXpress(v, width = 672, height = 500)
```

## Horizontal Error Bars
```{r}
f <- ggplot(
  df.summary, 
  aes(x = len, y = dose, xmin = len-sd, xmax = len+sd)
  )
h <- f + geom_point() + geom_errorbarh(height=.2)
h
canvasXpress(h, width = 672, height = 500)
```

### Senssion Info
```{r}
sessionInfo()
```

