---
title: "Scaling Axes"
author: "Isaac Neuhaus"
date: "06/04/2024"
output:
  html_document:
    toc: yes
    theme: cerulean
    smart: no
    self_contained: yes
editor_options: 
  chunk_output_type: console
---

# Load data and Libraries
### https://r4ds.had.co.nz/data-visualisation.html
```{r setup}
if (!("ggplot2" %in% installed.packages())) {
    install.packages('ggplot2')
}
library('ggplot2')
if (!("devtools" %in% installed.packages())) {
    install.packages('devtools')
}
if (!("canvasXpress" %in% installed.packages())) {
    devtools::install_github('neuhausi/canvasXpress')
}
#devtools::install_local("~/git/canvas/R/canvasXpress.tar.gz", build_manual = TRUE, upgrade = "always")
library('canvasXpress')
```


## Box plot
```{r}
bp <- ggplot(ToothGrowth, aes(x=factor(dose), y=len, group = dose)) + geom_boxplot()
bp
canvasXpress(bp)
```

## scatter plot
```{r}
sp<-ggplot(cars, aes(x = speed, y = dist)) + geom_point()
sp
canvasXpress(sp)
```

## Box plot : change y axis range
```{r}
bp2 <- bp + ylim(0,50)
bp2
canvasXpress(bp2)
```

## scatter plots : change x and y limits
```{r}
sp2 <- sp + xlim(5, 40) + ylim(0, 150)
sp2
canvasXpress(sp2)
```

## Change x and y axis labels, and limits
```{r}
sp3 <- sp + scale_x_continuous(name="Speed of cars", limits=c(0, 30)) + scale_y_continuous(name="Stopping distance", limits=c(0, 150))
sp3
canvasXpress(sp3)
```

## Default scatter plot
```{r}
sp4 <- ggplot(cars, aes(x = speed, y = dist)) + geom_point()
sp4
canvasXpress(sp4)
```

## Log transformation using scale_xx() possible values for trans : 'log2', 'log10','sqrt'
```{r}
sp5 <- sp4 + scale_x_continuous(trans='log2') + scale_y_continuous(trans='log2')
sp5
canvasXpress(sp5)
```

## Sqrt transformation
```{r}
sp6 <- sp4 + scale_y_sqrt()
sp6
canvasXpress(sp6)
```

## Reverse coordinates
```{r}
sp7 <- sp4 + scale_y_reverse() 
sp7
canvasXpress(sp7)
```

## Possible values for x and y : "log2", "log10", "sqrt", ...
```{r}
sp8 <- sp4 + coord_trans(x="log2", y="log2")
sp8
canvasXpress(sp8)
```

## log-log plot without log tick marks
```{r}
library(MASS) # to access Animals data sets
library(scales) # to access break formatting functions
# x and y axis are transformed and formatted
p2 <- ggplot(Animals, aes(x = body, y = brain)) + geom_point() +
     scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
     scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
     theme_bw()
p2
canvasXpress(p2)
```

## Show log tick marks
```{r}
p3 <- p2 + annotation_logticks()
p3
canvasXpress(p3)
```
