---
title: "CanvasXpress - Seurat"
author: "Isaac Neuhaus"
date: "06/04/2024"
output:
  html_document:
    toc: yes
    theme: cerulean
    smart: no
    self_contained: yes
---


```{r setup}
if (!("ggplot2" %in% installed.packages())) {
    install.packages('ggplot2')
}
library('ggplot2')
if (!("Seurat" %in% installed.packages())) {
    install.packages('Seurat')
}
library('Seurat')
if (!("devtools" %in% installed.packages())) {
    install.packages('devtools')
}
if (!("canvasXpress" %in% installed.packages())) {
    devtools::install_github('neuhausi/canvasXpress')
}
#devtools::install_local("~/git/canvas/R/canvasXpress.tar.gz", build_manual = TRUE, upgrade = "always")
library('canvasXpress')
```

# Load data and Libraries
### https://satijalab.org/seurat/articles/visualization_vignette.html
```{r}
SeuratData::InstallData("pbmc3k")
```

```{r}
library(Seurat)
library(SeuratData)
library(ggplot2)
library(ggridges)
library(patchwork)
pbmc3k.final <- LoadData("pbmc3k", type = "pbmc3k.final")
pbmc3k.final$groups <- sample(c("group1", "group2"), size = ncol(pbmc3k.final), replace = TRUE)
features <- c("LYZ", "CCL5", "IL32", "PTPRCAP", "FCGR3A", "PF4")
pbmc3k.final
```
# Five visualizations of marker feature expression

### Ridge plots - from ggridges. Visualize single cell expression distributions in each cluster
```{r}
r = RidgePlot(pbmc3k.final, features = features, ncol = 2)
r
canvasXpress(r)
```

### Dot plots - SplitDotPlotGG has been replaced with the `split.by` parameter for DotPlot
```{r}
d = DotPlot(pbmc3k.final, features = features, split.by = "groups") + RotatedAxis()
d
canvasXpress(d)
```

### Violin plot - Visualize single cell expression distributions in each cluster
```{r}
v = VlnPlot(pbmc3k.final, features = features)
v
canvasXpress(v)
```

### Violin plots can also be split on some variable. Simply add the splitting variable to object metadata and pass it to the split.by argument
```{r}
v2 = VlnPlot(pbmc3k.final, features = "percent.mt", split.by = "groups")
v2
canvasXpress(v2)
```

### Feature plot - visualize feature expression in low-dimensional space
```{r}
f = FeaturePlot(pbmc3k.final, features = features)
f
canvasXpress(f)
```

### DimPlot replaces TSNEPlot, PCAPlot, etc. In addition, it will plot either 'umap', 'tsne', or 'pca' by default, in that order
```{r}
s = DimPlot(pbmc3k.final)
s
canvasXpress(s)
```

### Dot plots - SplitDotPlotGG has been replaced with the `split.by` parameter for DotPlot
```{r}
d = DotPlot(pbmc3k.final, features = features, split.by = "groups") + RotatedAxis()
d
canvasXpress(d)
```

### Single cell heatmap of feature expression
```{r}
h = DoHeatmap(subset(pbmc3k.final, downsample = 100), features = features, size = 3)
h
canvasXpress(h)
```

### Senssion Info
```{r}
sessionInfo()
```

